Software Clusterings with Vector Semantics and the Call Graph

### 2019-06-29 manuscript/manuscript.tex
Finalize paper for camera-ready submission

In this paper, we propose a novel method to determine a software's modules without knowledge of its architectural structure, and empirically validate the method's performance. We cluster files by combining document embeddings, generated with the  \emph{Doc2Vec} algorithm, and the \emph{call graph}, provided by Static Graph Analyzers to an augmented graph. We use the \emph{Louvain Algorithm} to determine its community structure and propose a module--level clustering. 
Our method performs better in terms of stability, authoritativeness 
, and extremity over other state-of-the-art clustering methods proposed in the literature and is able to decently 
recover the ground truth clustering of the Linux Kernel. 
Finally, we conclude that semantic information from vector semantics as well as the call graph can produce 
accurate results for software clusterings of large systems.

@@

The codebase is initially preprocessed and the resulting data is fed into a Skip-Gram model for the initial components, which correspond to nodes in the call graph. Next, we add weights between the nodes with the normalized similarity between them. Finally, we run the Louvain algorithm to obtain a software clustering. The overall process is described below.

First of all, we tokenize the code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming \cite{wordninja} and lemmatize 
each individual sub-token via using the \textsc{nlp} package spaCy \cite{spacy}. 
For example, the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.
We, then, train a Doc2Vec model using Gensim \cite{gensim}. 
The codebase is also processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \emph{CScout} \cite{cscout} in order to generate the directed call graph through function calls between the files. 
Each source code file is assigned to a module. 
Modules can be user-defined or automatically generated by their respective directories at a desired directory tree depth. 
The input to the clustering procedure
is the call graph with cosine similarities of the modules as edge weights.
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
In case we want to consider edge directionality, we do a bipartite transformation, detect communities and
merge the results with a disjoint-set data structure as stated in reference \cite{malliaros}.

@@

We have chosen to use Linux as a codebase to evaluate our method on since it is a large and complex system with 27 years of continuous development and spanning 20.3 millions source lines of code (\textsc{sloc}) at the time of writing.  
Moreover, its ground truth is is easy to due to the clear directory structure and construct test-cases for
evaluations. 
Finally, it was also used in related studies \cite{acdc, evaluation} as a reference system for evaluation.

 Our method was tested on Linux 4.21,
consisting of 20.3 million \textsc{sloc} against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} using the same document embeddings
as well as \textsc{acdc} with structural information \cite{acdc} and \textsc{limbo} \cite{limbo} with Bag-of-Words features. The results are presented in Table \ref{tab:evaluation} and user-defined parameters are formatted in italics. 
As ground truth, we have used the first level directories as a target clustering and as input, we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. Since Louvain Community Detection produces different (but very similar) results from every time it runs, we ran the simulations 
multiple times and averaged the results. 
The experimentation with a large and complex system may constitute a threat to our findings' validity since the results are not general.

>>>>

In this paper, we propose a novel method to determine a software's modules without knowledge of its architectural structure, and empirically validate the method's performance. We cluster files by combining document embeddings, generated with the  \emph{Doc2Vec} algorithm, and the \emph{call graph}, provided by Static Graph Analyzers to an augmented graph. We use the \emph{Louvain Algorithm} to determine its community structure and propose a module-level clustering. 
Our method performs better in terms of stability, authoritativeness 
, and extremity over other state-of-the-art clustering methods proposed in the literature and is able to decently 
recover the ground truth clustering of the Linux Kernel. 
Finally, we conclude that semantic information from vector semantics as well as the call graph can produce 
accurate results for software clusterings of large systems.

@@

The codebase is initially preprocessed and the resulting data is fed into a Skip-Gram model for the initial components, which correspond to nodes in the call graph. Next, we add weights between the nodes with the normalized similarity $w(i, j) = \left (1 + \cos(\mathbf d_i, \mathbf d_j) \right ) / 2$ between the embeddings. Finally, we run the Louvain algorithm to obtain a clustering. The overall process is described below.

First of all, we tokenize the source code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming \cite{wordninja} and lemmatize 
each individual sub-token via using the \textsc{nlp} package spaCy \cite{spacy}. 
For example, the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.
We, then, train a Doc2Vec model using Gensim \cite{gensim}. 
The codebase is also processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \emph{CScout} \cite{cscout} in order to generate the directed call graph through function calls between the files. 
Each source code file is assigned to a module. 
Modules can be user-defined or automatically generated by their respective directories at a desired directory tree depth. 
The input to the clustering procedure
is the call graph with normalized cosine similarities of the modules as edge weights.
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
In case we want to consider edge directionality, we do a bipartite transformation, detect communities and
merge the results with a union-find data structure as stated in reference \cite{malliaros}.

@@

We have chosen to use Linux as a codebase to evaluate our method on since it is a large and complex system with 27 years of continuous development and spanning 20.3 millions source lines of code (\textsc{sloc}) at the time of writing.  
Moreover, it is easy to establish a ground truth due to its clear
structure and construct test-cases for
evaluations. 
Finally, it was also used in related studies \cite{acdc, evaluation} as a reference system for evaluation.

 Our method was tested on Linux 4.21,
consisting of 20.3 million \textsc{sloc} against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} using the same document embeddings
as well as \textsc{acdc} with structural information \cite{acdc} and \textsc{limbo} \cite{limbo} with Bag-of-Words features. The results are presented in Table \ref{tab:evaluation} and user-defined parameters are formatted in italics. 
As ground truth, we have used the first level directories as a target clustering and as input, we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. Since Louvain Community Detection produces different, but very similar, results from every time it runs, we ran the simulations 
multiple times and averaged the results. 
The experimentation with a large and complex system may constitute a threat to our findings' validity since the results are not general.
